---
title: "Likelihood (and Seals)"
author: "STAT 245"
date: ""
output:
  xaringan::moon_reader:
    css: ['default', 'https://stackpath.bootstrapcdn.com/bootstrap/4.3.1/css/bootstrap.min.css', 'slides.css']
    lib_dir: libs
    nature:
      titleSlideClass: ['inverse','middle','left',my-title-slide]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---

```{r, setup, include=FALSE}
require(tidyverse)
require(mosaic)    
require(ggformula)
knitr::opts_chunk$set(echo = TRUE, 
                      fig.width = 7, 
                      fig.height = 3,
                      tidy = FALSE,
                      fig.align = 'center', 
                      message = FALSE, 
                      warning = FALSE,
                      error = TRUE,
                      out.width = '60%', 
                      dpi = 300)
theme_set(theme_minimal(base_size = 22))

set.seed(9)
harp_seals <- data.frame(ice_cover = runif(18, min = 10, max = 80)) |>
  mutate(strandings = 75.22 - 0.88*ice_cover + rnorm(18, 0, 16)) |>
  mutate(strandings = ifelse(strandings < 0, -strandings, strandings))

```


# Harp Seal Strandings
### Johnston *et al.,* ["Effects of Climate Change on Harp Seals"](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0029158#s4)
.pull-left[
.small[Harp seals use seasonal sea ice as a place to breed. Does climate change affect their survival? How many young-of-year are found stranded (dead) annually?]
]

.pull-right[
```{r, out.width = '100%', echo = FALSE}
knitr::include_graphics('https://encrypted-tbn3.gstatic.com/licensed-image?q=tbn:ANd9GcQwWjz3zKdh83FpMyKqJ2FgcZFceYZlK5eRyVlDal4VZo3iMSFUzcOYp9TmTH2c-waEG-4V5QJXB-OaWEI')
```

.smaller[Tier Und Naturfotografie J und C Sohns / Getty Images]
]
---

# Annual Strandings

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
gf_dhistogram(~strandings, data = harp_seals,
              binwidth = 20) |>
  gf_lims(x = c(-25, 115)) |>
  gf_labs(x='Young-of-Year Stranded Harp Seals', 
          y='Probability\nDensity') |>
  gf_fitdistr(dist = 'dnorm', size = 1.3) 
```

---

# Normal probability density function (PDF)
$$ f(x) = \frac{1}{\sqrt{2\pi\sigma^2}} e^{-\frac{(x-\mu)^2}{2\sigma^2}} $$

---

# A simple model

- Stranding distribution description:
<br><br>

```{r}
df_stats(~strandings, data = harp_seals, mean, sd) 
```

---

# Model-based Predictions

.small[
- Guess the number of strandings that will happen this year.
- Is it more likely that there will be 40 or 80 strandings? 
- How much more likely?
- What is the *probability* of there being 47 strandings this year?
]

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=2.5}
gf_dhistogram(~strandings, data = harp_seals,
              binwidth = 20) |>
  gf_lims(x = c(-25, 115)) |>
  gf_labs(x='Young-of-Year Stranded Harp Seals', 
          y='Probability\nDensity') |>
  gf_fitdistr(dist = 'dnorm', size = 1.3) 
```

---

# Likelihood to the Rescue!
## Which is **more likely?**

- **Three years with 36, 41, and 43 strandings**
- Three years with 60, 41, and 40 strandings

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=2.5}
gf_dhistogram(~strandings, data = harp_seals,
              binwidth = 20) |>
  gf_lims(x = c(-25, 115)) |>
  gf_labs(x='Young-of-Year Stranded Harp Seals', 
          y='Probability\nDensity') |>
  gf_fitdistr(dist = 'dnorm', size = 1.3) 
```

---

# Likelihood approach
- How did you:
  - Find the likelihood of each observation?
  <br>
  - Combine the likelihoods of a set of three observations?
  <br>
  - What did you have to assume about the set of observations?
  
---
class: huge center middle inverse subsection

# How does all this seal stuff relate to linear regression?

---
# Likelihood and `lm()`
## How can we rewrite our model as a linear regression?

```{r}
lm_version <- lm(strandings ~ 1, 
                 data = harp_seals)
```

---
class: small

# Model Equation

```{r}
summary(lm_version)
```

---

# Compare models w/Likelihood
## it's likelihood *of the data, given a particular model*

---

# Compare Models
## w/Likelihood - Class example (teamwork problem)

```{r}
gf_point(strandings ~ ice_cover,
         data = harp_seals)
```
---

```{r}
lm_version_2 <- lm(strandings ~ ice_cover, 
                           data = harp_seals)
msummary(lm_version_2)
```

---
## Model Comparison Challenge
### *joint likelihood* of residuals given $\sigma$

```{r}
harp_seals <- harp_seals |>
  mutate(resid1 = resid(lm_version),
         resid2 = resid(lm_version_2))

```

```{r, echo = FALSE}
harp_seals <- harp_seals |>
  mutate(across(where(is.numeric), round, digits = 1))
write_csv(harp_seals, 'harp_seal_likelihood.csv')
```

<https://cutt.ly/seal-likelihood>
.small[<https://www.danielsoper.com/statcalc/calculator.aspx?id=54>]

---

# Likelihood...

- Can be used to measure model-data match
- (...and then as ingredient to AIC/BIC)
- What about *other* probability distributions?
- [Next...] Can be used to *fit* one model: which parameter estimates are "best"?
