---
title: "Linear Model Equation Practice"
format: 
  pdf: default
  html:
    embed-resources: true
    format-links: true
editor: visual
---

Metadata available at: <https://stacyderuiter.github.io/stat245-fa24/tehran-glucose-meta.html>

```{r}
#| label: setup
#| include: false

library(tidyverse)

TGS <- read_csv('https://sldr.netlify.app/data/tgs.csv')
```

```{r}
chol_model <- lm(total_cholesterol ~ age,
                 data = TGS)
summary(chol_model)
```

1.  What are the predictor and response variables?
2.  What question can this model answer?
3.  Write the model equation, using information from the model
    `summary()`.

\newpage

```{r}
bp_model <- lm(SBP ~ triglyceride, 
               data = TGS)
summary(bp_model)
```

1.  What are the predictor and response variables?
2.  What question can this model answer?
3.  Write the model equation, using information from the model
    `summary()`.

\newpage

```{r}
tri_model <- lm(triglyceride ~ age, 
               data = TGS)
summary(tri_model)
```

1.  What are the predictor and response variables?
2.  What question can this model answer?
3.  Write the model equation, using information from the model
    `summary()`.

\newpage

```{r}
tri_model2 <- lm(triglyceride ~ total_cholesterol, 
               data = TGS)
summary(tri_model2)
```

1.  What are the predictor and response variables?
2.  What question can this model answer?
3.  Write the model equation, using information from the model
    `summary()`.

\newpage

```{r}
bmi_model <- lm(BMI ~ age, 
               data = TGS)
summary(bmi_model)
```

1.  What are the predictor and response variables?
2.  What question can this model answer?
3.  Write the model equation, using information from the model
    `summary()`.

\newpage

```{r}
whr_model <- lm(SBP ~ WHR100, 
               data = TGS)
summary(whr_model)
```

1.  What are the predictor and response variables?
2.  What question can this model answer?
3.  Write the model equation, using information from the model
    `summary()`.
