---
title: "What's your model process map?"
author: "STAT 245"
date: ""
output:
  xaringan::moon_reader:
    css: ['default', 'https://stackpath.bootstrapcdn.com/bootstrap/4.3.1/css/bootstrap.min.css', 'slides.css']
    lib_dir: libs
    nature:
      titleSlideClass: ['inverse','middle','left',my-title-slide]
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%total%"
      ratio: '16:9'
---

```{r, setup, include=FALSE}
library(tidyverse)
library(mosaic)    
library(ggformula)
knitr::opts_chunk$set(echo = TRUE, 
                      fig.width = 7, 
                      fig.height = 3,
                      tidy = FALSE,
                      fig.align = 'center', 
                      message = FALSE, 
                      warning = FALSE,
                      error = TRUE,
                      out.width = '60%', 
                      dpi = 300)
theme_set(theme_minimal(base_size = 22))

```

```{r, include=FALSE}
sscrime <- read_csv('https://sldr.netlify.app/data/sscrime.csv') |>
  mutate(Location = fct_relevel(Location, 'City', 'Town', 'Urban Fringe', 'Rural'))

theft_pois <- glm(Thefts ~ NEnrollment + Location + 
                 TrainingHours + SecurityCameras, 
               data = sscrime,
               family = poisson(link = 'log'))
```



class: huge inverse center middle subsection
# Response Type

---

class: huge inverse center middle subsection
# Model Family & Link Function

---

class: huge inverse center middle subsection
# Assessment Conditions

---

class: huge inverse center middle subsection
# Check Using Plots:

---

# Multiple Linear Regression

---



# Poisson Regression

---


# Negative Binomial Regression

---



# Binary Regression

---

# Lack of non-linearity

---

# Independence of Residuals

---

# Normality of Residuals

---

# Constant Residual Variance

---

# Mean-Variance Relationship as expected

---

# Logit-Linearity (or probit- or cloglog-linearity)

---

# Log-linearity

---

# Residuals vs fitted values

---
# Residual histogram

---
# ACF plot

---
# Scaled residual plot

---
# Scatter plot of response (or log(response) or logit(response) etc.) vs quantitative predictor(s)

---
# Quantitative variable

---
# Count variable

---

# Categorical variable (2 categories)

---

# Categorical variable (3+ categories)

---

# Counts of "successes" and "failures"


