---
title: "This is your project title"
author: "Student and partner names go here"
format: 
  html:
    embed-resources: true
    code-tools: true
    theme: lux
    page-layout: custom
    fontsize: '80%'
editor: visual
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(ggformula)
library(ggforce)

knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      fig.width = 7,
                      fig.height = 5
                      )
theme_set(theme_minimal(base_size = 20))
```

<!-- style notes... -->
<!-- we used the "lux" theme - see more at: -->
<!-- https://quarto.org/docs/output-formats/html-themes.html -->
<!-- consider more customization of colors, etc. - some options at
<!-- https://quarto.org/docs/output-formats/html-themes.html -->
<!-- or use css if you know how -->


::: {.grid}

::: {.g-col-4}
<!-- This column takes 1/3 of the page -->
# Introduction

- Text or bullet points giving context
  - Describe your partner organization, 
  - your data, and 
  - the question(s) you're trying to answer. 
- **Refer here to the bibliographic resources you named in your Statement of Work!**

# Data
- Present key graphs of your data 
  - Include text (or titles!) clarifying the main point of each graph. 
  - Graphs should score well on our [graphics critique checklist](https://moodle.calvin.edu/pluginfile.php/2373925/mod_page/intro/Graphics%20Critique%20Checklist.pdf)!
- Consider replacing **Data** with a more-informative header...

```{r, echo = FALSE, out.width='95%'}
gf_boxplot(mpg ~ factor(cyl), data = mtcars,
        color = ~(cyl==8)) |>
  gf_sina() |>
  gf_labs(x = 'Cylinders',
          y = 'Miles per Gallon',
          title = '8-Cylider cars guzzle gas') |>
  gf_refine(scale_color_brewer('',
                               palette = 'Set2')) |>
  gf_theme(legend.position = 'none') 
```
:::

::: {.g-col-4}
<!-- This column takes 1/3 of the page -->
# Statistical Modeling

- Code-free description of the model you fitted. 
  - Short! 
  - In words! 
  - See the [literature papers you worked on](https://stacyderuiter.github.io/stat245-sp24/re-lit-examples.html) for inspiration (Additional examples will be given.)

# Main Findings

- Interpretation of model results. What are the answers to your research question(s)? 
- You might briefly refer to model selection results or assessment but they should probably not be shown in detail. 
- You may show prediction plot(s) or you may just refer back to graphs of the data
:::

::: {.g-col-4}
<!-- This column takes 1/3 of the page -->
# Your experience

- What were the biggest challenges and triumphs working on this project?
  - These might be technical, or not. 
  - Include here the kind of things you'd say if you were describing what you learned from this experience as part of a job interview, 
  - or giving advice to a student about to take STAT 245 in the future...

# References & Acknowledgements

- Thank your partner?
- Include a bibliography (any consistent style is ok; here is [a guide for MLA style.](https://owl.purdue.edu/owl/research_and_citation/mla_style/mla_formatting_and_style_guide/mla_formatting_and_style_guide.html))

## Works Cited 

::: {style="font-size: 70%;"}
Alford, K. R., L. K. Walls, S. L. DeRuiter, R Pruim, J. E. VanHorn, M. Bone, J. Deighton, J. Koeman, S. Kornoelje, A. Koroma, R. LeTourneau, B. Lika, K. Rosendale, and N. Wang. 2023. "Reduction in Diarrhoea Rates after Household Water Filter Distribution in Small and Remote Communities in Liberia." Cogent Public Health 10, no. 1. doi:10.1080/27707571.2023.2205716.
:::

:::

:::