---
title: "This is your project title"
author: "Student and partner names go here"
format: 
  html:
    embed-resources: true
    code-tools: true
    theme: lux
    page-layout: custom
    fontsize: '80%'
editor: visual
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(ggformula)
library(ggforce)

knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      fig.width = 7,
                      fig.height = 4
                      )
theme_set(theme_minimal(base_size = 20))
```

<!-- style notes... -->

<!-- we used the "lux" theme - see more at: -->

<!-- https://quarto.org/docs/output-formats/html-themes.html -->

<!-- consider more customization of colors, etc. - some options at
<!-- https://quarto.org/docs/output-formats/html-themes.html -->

<!-- or use css if you know how -->

::::::: grid
::: g-col-4
<!-- This column takes 1/3 of the page -->

# Introduction

-   Text or bullet points giving context:
    -   Describe your partner organization,
    -   your data, and
    -   the question(s) you're trying to answer.
    -   Why is *your partner's mission* and *the specific question
        you're addressing* important?
    -   What previous similar work has been done?
-   **You** *must* r**efer here to 2+ bibliographic resources you
    identified in your Statement of Work! Use (Author, Year) in-text
    citations.**
-   This is the only section that might actually be a longer,
    carefully-written block of text.

# Data

-   Present key graphs of your data
-   See Class Resources on Moodle 12/3 for saving data to read in here,
    or saving image files of graphs. For poster, use more titles and
    bigger fonts/symbols!
    -   Include text (or titles!) clarifying the main point of each
        graph.
    -   Graphs should score well on our [graphics critique
        checklist](https://moodle.calvin.edu/pluginfile.php/2522988/mod_label/intro/Graphics%20Critique%20Checklist.pdf)!
-   I recommend: replacing **Data** with a more-informative header...

```{r, echo = FALSE, out.width='95%'}
gf_boxplot(mpg ~ factor(cyl), data = mtcars,
        color = ~(cyl==8)) |>
  gf_sina() |>
  gf_labs(x = 'Cylinders',
          y = 'Miles per Gallon',
          title = '8-Cylider cars guzzle gas',
          caption = "Figure 1. Irrelevant boxpots about cars.\nCaptions like this can probably be\nOMITTED if you have good titles and labels!") |>
  gf_refine(scale_color_brewer('',
                               palette = 'Set2')) |>
  gf_theme(legend.position = 'none') 
```
:::

::: g-col-4
<!-- This column takes 1/3 of the page -->

# Statistical Modeling

-   Code-free description of the model you fitted.
    -   Short! In words!
    -   Make it clear what model type(s), response(s), predictors, RE
        were used.
    -   See the [literature papers you worked
        on](https://stacyderuiter.github.io/stat245-sp24/re-lit-examples.html)
        for inspiration (Additional examples will be given.)

# Main Findings

-   Interpretation of model results. What are the answers to your
    research question(s)?
-   You might briefly refer to model selection results or assessment but
    they should probably not be shown in detail.
-   You may show prediction plot(s), but if possible, rely on graphs of
    the data
:::

:::: g-col-4
<!-- This column takes 1/3 of the page -->

# Your experience

-   What were the biggest challenges and triumphs working on this
    project?
    -   These might be technical, or not.
    -   Include here the kind of things you'd say if you were describing
        what you learned from this experience as part of a job
        interview,
    -   or giving advice to a student about to take STAT 245 in the
        future...

# References & Acknowledgements

-   Thank your partner?
-   Include a bibliography (any consistent style is ok; here is [a guide
    for MLA
    style.](https://owl.purdue.edu/owl/research_and_citation/mla_style/mla_formatting_and_style_guide/mla_formatting_and_style_guide.html))

## Works Cited

*see above for link to MLA citation style format examples* - *include
all sources cited on poster.*

::: {style="font-size: 70%;"}
Alford, K. R., L. K. Walls, S. L. DeRuiter, R Pruim, J. E. VanHorn, M.
Bone, J. Deighton, J. Koeman, S. Kornoelje, A. Koroma, R. LeTourneau, B.
Lika, K. Rosendale, and N. Wang. 2023. "Reduction in Diarrhoea Rates
after Household Water Filter Distribution in Small and Remote
Communities in Liberia." Cogent Public Health 10, no. 1.
doi:10.1080/27707571.2023.2205716.
:::
::::
:::::::
