---
title: "Fall 2024 STAT 245 Project Teams"
format: 
  html:
    embed-resources: true
editor: visual
---

```{r}
#| include: false
#| label: setup

library(tidyverse)
```

## Information

In creating project teams, I balance preferred projects and teammates
and schedules, working as hard as I could to make sure all teams have
some outside-class meeting times available each week.

I also have to make that every project is staffed, and that teams are
balanced in terms of personalities and strengths (as best I know them so
early in the semester). Please know it's a complex balancing act and I
did my best to maximize all teams' chances of success and happiness.

**Please check your team assignment and reach out to me right away in
case of a mistake or problem. I will do my best to address any
concerns.**

## What's Next?

1.  Around mid-day Monday Sept. 23**, I will send an email to introduce
    each team** to their partner organization/contact person.

2.  **Before emailing the partner/contact person**, student teams should
    communicate among themselves to make some initial decisions. You
    will have time to confer in class on 9/24 but that would be
    wonderful if you confer via email etc. Monday. To help get started,
    emails are included in the table below.

    -   **Elect a point-of-contact**: Choose one team member who is
        willing, organized, responsible, and comfortable writing
        professional emails. ***They will be the one student responsible
        for email communication with the partner.*** (This avoids
        confusion - the partner knows who to contact and won't ever get
        redundant or contradictory messages from different students.)

    -   **Identify some possible meeting times** - ideally during
        business hours - when you may be able to have your initial
        meeting with your partner.

        -   Target meeting during the week of Sept 23-27 or Sept 30 -
            Oct 4. (Hopefully it won't be later than this, but if your
            partners travel/work schedule requires a delay, just keep
            Prof DR informed.)

        -   *At this first meeting, you will learn about your partner's
            work, as well as hearing about project background and goals.
            Try to meet in person if feasible, either on campus or in
            town. If not, a videoconference is ok.*

        -   *Do your best to accommodate your partner org's busy
            schedule! You will likely only meet with them in-person
            twice, at the start and end of the project, so do your very
            best to ensure everyone can attend.*

        -   *Prof DR will not attend these (unless she's the partner) -
            you are in charge!*

3.  Once you have a point-of-contact student and some proposed times
    identified, have the point-of-contact student reply to the partner
    to introduce themselves and propose/arrange meeting times.

4.  Also try to block out one or more weekly meeting times you can all
    commit to. (Plan to meet weekly and you can always cancel if a
    meeting really isn't needed some week.)

5.  You will get more information in class on Sept 24+. If you somehow
    arrange to meet with your partner *before* the 24th and have
    questions*,* please let Prof DR know so she can share additional
    info and advice.

## Teams

```{r}
#| echo: false
#| label: team-table

teamz <- readr::read_csv('project-teams.csv', show_col_types = FALSE) |>
  mutate(Name = paste(`First name`, `Last name`)) |>
  rename(Email = `Email address`) |>
  select(Name, Project, Email) |> 
  mutate(Email = stringr::str_remove_all(Email, "@calvin.edu")) |>
  drop_na(Project) |> arrange(Project)

DT::datatable(teamz,
              caption = NULL,
              options = list(dom = 'Blfrtip',
                             lengthMenu = list(c(-1, 10, 15),
                                               c('All', '10', '15')),
                             paging = T))
```
