---
title: "Saving Prepared Data"
format: 
  html:
    embed-resources: true
    code-tools: true
    toc: true
editor: visual
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
knitr::opts_chunk$set(eval = FALSE)
```


# Motivation
What if you want to have one Quarto document in which you do all your data prep...

And then all your graphics and analysis in a second, tech report document?

Or, what if you want separate report documents for several different research questions that all use the same dataset?

Or, what if you want to save a .csv file with your prepared, clean dataset to share with your project partner?

The approach below will allow you to save the prepped dataset in one of two formats:

- **an .RMD file** (good for reading in in another quarto document, since things like variable types and factor level order will be preserved; can really only open in R)
- **a .csv file** (good for sharing with partner; easy to open without using R)

# Example Data

Here, we will read in and "prepare" an example dataset. Instead of this code, you'll have *all your existing code to read in your project data from its original format and then clean and prepare it.*

```{r}
FSSA_data <- read_csv('https://sldr.netlify.app/data/indiana_fssa.csv',
                      show_col_types = FALSE) |>
  mutate(TANF_payments_millions = TANF_payments / 1e6,
         TANF_families_10k = TANF_families / 10000,
         COVID = fct_relevel(COVID, 'pre-pandemic', 'pandemic'),
         month = fct_relevel(month, "January", "February", "March", "April",
                             "May", "June", "July", "August",
                             "September", "October", "November", "December"))
```

# Saving RDS and csv files

Now, to save the fully-prepared dataset `FSSA_data` to a file called FSSA_data.RDS, we run:

```{r}
saveRDS(FSSA_data, file = "FSSA_data.RDS")
```

If we want to also (or instead) save a .csv instead, the code would be:

```{r}
write_csv(FSSA_data, file = "FSSA_data.csv")
```

# Reading .RDS files

We can read in the fully-prepped dataset that we saved as an RDS file by running code like this in any qmd:

```{r}
FSSA_data <- readRDS("FSSA_data.RDS")
```

(So you can have multiple qmd files as part of your project work, *without* copy-pasting or re-running all your data prep code!)



